---
title: "Preparation_Amazon_data"
date: "2022-12-28"
output:
  html_document: 
    number_sections: yes
    fig_caption: yes
    toc: true
    toc_float: 
      collapsed: true
    theme: cerulean
    highlight: kate
    toc_depth: 5
    keep_md: yes
    df_print: paged
---

# Load libraries

```{r}
library(tictoc)
library(tidyverse)
library(dplyr)
library(terra)
library(raster)
library(sf)
library(RColorBrewer)
library(scico)
library(tidyterra)
library(ggforce)
library(latex2exp)
library(rts)
library(doParallel)
library(parallel)
library(foreach)
library(data.table)
```

# Functions

```{r}
#---- Function to rename layers ----
renameLayers <- function(dataRast, fileStart, prefix){
  # index of file name in the path
  id.date <- unlist(gregexpr(fileStart, sources(dataRast)[[1]])) + nchar(fileStart)
  # rename layers
  names(dataRast) <- 
    substr(sources(dataRast), id.date, (id.date+50)) %>%
    gsub(".tif", "_1", .) %>%
    as.Date("%Y_%m_%d") %>%
    format(., '%Y_%m') %>%
    paste0(prefix, .)
  
  return(dataRast)
}

#---- Plot Function ----
myPlot <- function(
  rast, 
  title=NULL, 
  sub_title=NULL, 
  theme=2, 
  xy.limit = list(xmin=NULL, xmax=NULL, ymin=NULL, ymax=NULL),
  xy.zoom = list(xmin=NULL, xmax=NULL, ymin=NULL, ymax=NULL, zoom.size=NULL),
  max_cell=1e8,
  x_angle=0,
  b_size=16,
  na.color=NA,
  v_unit=c(2, 2, 2, 2) # c(top, right , below , left)
){
  p1 <- ggplot()
  p1 <- switch(
    theme,
    p1,
    p1 + theme_bw(base_size=b_size),
    p1 + theme_linedraw(base_size=b_size),
    p1 + theme_light(base_size=b_size),
    p1 + theme_minimal(base_size=b_size),
    p1 + theme_classic(base_size=b_size),
    p1 + theme_gray(base_size=b_size),
    p1 + theme_dark(base_size=b_size)
  )
  p1 <- p1  +
    geom_spatvector(data = amaz.basin.shp$geometry, fill = na.color, color = "gray40") +
    geom_spatraster(data = rast, maxcell = max_cell) +
    scale_x_continuous(limits = c(xy.limit$xmin, xy.limit$xmax), 
                       labels = function(x) format(x, scientific = T)) +
    scale_y_continuous(limits = c(xy.limit$ymin, xy.limit$ymax), 
                       labels = function(x) format(x, scientific = T)) +
    ggtitle(label=title, subtitle=sub_title) +
    coord_sf(datum = pull_crs(rast)) +
    theme(
      axis.text.x = element_text(angle = x_angle)
      , plot.margin = unit(v_unit, "pt")
    )
  
  if (!all(sapply(xy.zoom, is.null))) {
    p1 <- p1 + facet_zoom(xlim = c(xy.zoom$xmin, xy.zoom$xmax),
                          ylim = c(xy.zoom$ymin, xy.zoom$ymax), 
                          zoom.size=xy.zoom$zoom) 
  }
  
  if (theme == 1) {p1 <- p1 + theme_void(base_size=b_size)}

  return(p1)
}
```

# Initialization

```{r}
my.path <- "~/Documents/"
path.data <- paste0(my.path, "/Amazon_new_data")
# Create a sequence date
seq.dates <- seq(as.Date("2001-1-1"), as.Date("2020-12-1"), by = "month")
# Create name of layers
ordered.names <- format(seq.dates, '%Y_%m')
# Import shape file
amaz.basin.shp <- st_read(
  paste0(path.data,"/0. Amazon_shapefile/projected/amazon_shp_projected.shp"))
```

#-

# Burnt Area data

## Import data

```{r}
# list of files
amaz.burntArea.list <- list.files(
  paste0(path.data,"/1. Burnt Area/03. Working Data"),
  full.names=TRUE,
  pattern = ".tif$")
# Import data with "Terra"
burntArea.rast <- rast(amaz.burntArea.list)
burntArea.rast
```

## Rename layers

```{r}
# Rename layers
burntArea.rast <- renameLayers(burntArea.rast, 'burntarea_working_', '')
burntArea.rast
```

## Order layers

```{r}
burntArea.rast <- burntArea.rast[[ordered.names]]
burntArea.rast
```

## Verification of the values

```{r}
# Verification of the values
burntArea.minmax <- minmax(burntArea.rast) %>% t() %>% as.data.frame()
burntArea.minmax[which((burntArea.minmax[,1] != -2) & (burntArea.minmax[,2] != 1)),]
```

## Plot the month of October 2020

```{r , warning=FALSE}
# Create the `rts` object
burntArea.rts <- rts(burntArea.rast, seq.dates)
# Applying the mask to plot only the amazon area.
ba <- burntArea.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# Change values as categorical 
levels(ba) <- data.frame(id=c(-2, 0, 1), val=c('-2', '0', '1'))
# Plot
my.colors <- c("mediumblue", "mediumseagreen", "firebrick")
p.ba <- myPlot(ba, title = "Burnt Area") +
  scale_fill_manual(name = NULL, values = my.colors, na.translate=FALSE) 
p.ba
```

```{r}
# Convert the raster object to a datatable
ba.dt <- as.data.table(ba, cell=T, xy=T)
# Plot
ggplot(data = ba.dt, aes(x = val)) + 
  geom_bar(stat = "count", aes(fill = val), position = "dodge") + 
  labs(title="Burnt Area in October 2020", x="Burnt Area") +
  coord_flip() + 
  scale_fill_manual(name = "Burnt Area", values = my.colors) +
  stat_count(geom = "text", 
             aes(label = ..count..),
             position=position_stack(vjust=0.5),
             colour = "black", size = 3.5) + 
  theme_bw(base_size=16)
```

## Percentage of fires

```{r}
freq.dt <- matrix(nrow = 0, ncol = 3) %>% as.data.table()
colnames(freq.dt) <- c("layer", "0", "1")
for (ras_id in amaz.burntArea.list){ # ras_id <- amaz.burntArea.list[1]
  cat("\n", ras_id)
  ras <- rast(ras_id) %>%
    renameLayers(., 'burntarea_working_', '') %>%
    mask(mask = amaz.basin.shp)
 
  # Replace -2 and -1 value by `NA`
  ras[ras %in% c(-2, -1)] <- NA
  ras.freq <- freq(ras, digits=0, usenames=T) %>% as.data.table()
  tmp <- dcast(ras.freq,layer ~ value,value.var = c("count"))
  freq.dt <- rbind(freq.dt, tmp)
}
percentage.fires <- sum(freq.dt[, '1']) / sum(freq.dt[, c('0', '1')])
```

## Missing Data

```{r}
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl, cores=detectCores() - 1)

# Raster to datatable in parallel: one raster per thread
rasList <- foreach (
  ras_id=amaz.burntArea.list, 
  .packages=c('terra', 'sf', 'dplyr'), 
  .combine='c') %dopar% {
    # Read and rename raster
    ras <- rast(ras_id) %>% renameLayers(., 'burntarea_working_', '')
    # Replace negative value by `NA`
    if (names(ras) %in% c("2012_07", "2012_09")) {ras[ras == -1] <- NA}
    # Count the missing data
    ras.freq.na <- not.na(ras) %>%
      mask(amaz.basin.shp) %>%
      terra::freq(., digits=0, value=0, usenames=T)
  
    list(ras.freq.na)
  }
stopCluster(cl)

# Bind all per-raster into one dataframe
burntArea.freq.na <- rbindlist(rasList, fill=T, use.names=T)
# Order
colnames(burntArea.freq.na)[3] <- "burntArea_na"
burntArea.freq.na <- burntArea.freq.na[order(burntArea.freq.na$layer)]
burntArea.freq.na
```

#-

# Land Cover data

## Import data

```{r}
# list of files
amaz.landCover.list <- list.files(
  paste0(path.data,"/2. Land Cover/03. Working Data"),
  full.names=TRUE,
  pattern = ".tif$")
# Import data with "Terra"
landCover.rast <- rast(amaz.landCover.list)
landCover.rast
```

## Rename and order layers

```{r}
# Rename layers
landCover.rast <- renameLayers(landCover.rast, 'landcover_working_', '')
# Order layers
landCover.rast <- landCover.rast[[ordered.names]]
landCover.rast
```

## Verification of the values

```{r}
# Verification of the values
landCover.minmax <- minmax(landCover.rast) %>% t() %>% as.data.frame()
landCover.minmax[
  which((!landCover.minmax[,1] %in% c(0:10)) & (!landCover.minmax[,2] %in% c(0:10))), ]
```

## Plot the month of October 2020

```{r}
# Create the `rts` object
landCover.rts <- rts(landCover.rast, seq.dates)
# Applying the mask to plot only the amazon area.
lc <- landCover.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)  
# Change values as categorical 
levels(lc) <- data.frame(id=0:10, val=as.character(c(0:10)))
# Plot
p.lc <- myPlot(lc, title = "Land Cover") + 
  scale_fill_hypso_d(
    name = NULL,
    palette = "colombia_hypso", 
    na.translate=FALSE)
p.lc
```

## Missing Data

```{r}
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl, cores=detectCores() - 1)

# Raster to datatable in parallel: one raster per thread
rasList <- foreach (
  ras_id=amaz.landCover.list, 
  .packages=c('terra', 'sf', 'dplyr'), 
  .combine='c') %dopar% {
    # Read and rename raster
    ras <- rast(ras_id) %>% renameLayers(., 'landcover_working_', '')
    # Count the missing data
    ras.freq.na <- not.na(ras) %>%
      mask(amaz.basin.shp) %>%
      terra::freq(., digits=0, value=0, usenames=T)
  
    list(ras.freq.na)
  }
stopCluster(cl)

# Bind all per-raster into one dataframe
landCover.freq.na <- rbindlist(rasList, fill=T, use.names=T)
# Order
colnames(landCover.freq.na)[3] <- "landCover_na"
landCover.freq.na <- landCover.freq.na[order(landCover.freq.na$layer)]
landCover.freq.na
```

#-

# Precipitation data

## Import data

```{r}
# list of files
amaz.precipitation.list <- list.files(
  paste0(path.data,"/3. Precipitation/03. Working Data"),
  full.names=TRUE,
  pattern = ".tif$")
# Import data with "Terra"
precipitation.rast <- rast(amaz.precipitation.list)
precipitation.rast
```

## Rename and order layers

```{r}
# Rename layers
precipitation.rast <- renameLayers(precipitation.rast, 'precipitation_working_', '')
# Order layers
precipitation.rast <- precipitation.rast[[ordered.names]]
precipitation.rast
```

## Verification of the values

```{r}
# Verification of the values
precipitation.minmax <- minmax(precipitation.rast) %>% t() %>% as.data.frame()
precipitation.minmax
```

## Plot the month of October 2020

```{r}
# Create a sequence date for 'rts' object
precipitation.rts <- rts(precipitation.rast, seq.dates)
# Applying the mask to plot only the amazon area.
prec <- precipitation.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# Plot
p.prec <- myPlot(prec, title = "Precipitation") + 
  scale_fill_scico(
    name = TeX(r"($\textit{(mm/hr)}$)"),
    palette = "lapaz", 
    direction = -1,
    trans = "pseudo_log",
    breaks = c(0,10,50,200,550),
    na.value = "transparent")
p.prec
```

## Missing Data

```{r}
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl, cores=detectCores() - 1)

# Raster to datatable in parallel: one raster per thread
rasList <- foreach (
  ras_id=amaz.precipitation.list, 
  .packages=c('terra', 'sf', 'dplyr'), 
  .combine='c') %dopar% {
    # Read and rename raster
    ras <- rast(ras_id) %>% renameLayers(., 'precipitation_working_', '')
    # Count the missing data
    ras.freq.na <- not.na(ras) %>%
      mask(amaz.basin.shp) %>%
      terra::freq(., digits=0, value=0, usenames=T)
  
    list(ras.freq.na)
  }
stopCluster(cl)

# Bind all per-raster into one dataframe
precipitation.freq.na <- rbindlist(rasList, fill=T, use.names=T)
# Order
colnames(precipitation.freq.na)[3] <- "precipitation_na"
precipitation.freq.na <- precipitation.freq.na[order(precipitation.freq.na$layer)]
precipitation.freq.na
```

### Zoom Plot

```{r}
# Applying the mask to plot only the amazon area.
prec <- precipitation.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# define the zoom area
prec.xy.zoom <- list(xmin=-0.385e+06, xmax=-0.37e+06, ymin=ymin(prec), ymax=1.64e+06, zoom=0.4)
# Plot
p.prec.na <- myPlot(
  prec, title = "Precipitation", 
  max_cell=1e7,
  x_angle=90,
  b_size=12,
  na.color="black",
  xy.zoom = prec.xy.zoom
) + 
  scale_fill_scico(
    name = TeX(r"($\textit{(mm/hr)}$)"),
    palette = "lapaz", 
    direction = -1,
    trans = "pseudo_log",
    breaks = c(0,10,50,200,550),
    na.value = "transparent")
p.prec.na
```

#-

# Soil Moisture data

## Import data

```{r}
# list of files
amaz.soilMoisture.list <- list.files(
  paste0(path.data,"/4. Soil Moisture/03. Working Data"),
  full.names=TRUE,
  pattern = ".tif$")
# Import data with "Terra"
soilMoisture.rast <- rast(amaz.soilMoisture.list)
soilMoisture.rast
```

## Rename and order layers

```{r}
# Rename layers
soilMoisture.rast <- renameLayers(soilMoisture.rast, 'soilmoisture_working_', '')
# Order layers
soilMoisture.rast <- soilMoisture.rast[[ordered.names]]
soilMoisture.rast
```

## Verification of the values

```{r}
# Verification of the values
soilMoisture.minmax <- minmax(soilMoisture.rast) %>% t() %>% as.data.frame()
soilMoisture.minmax
```

## Frequency of the negative values

```{r}
soilMoisture.freq <- freq(soilMoisture.rast, digits=3, usenames=T)
soilMoisture.freq[soilMoisture.freq$value < 0,]
```

### Zoom Plot

```{r}
# Create the `rts` object
soilMoisture.rts <- rts(soilMoisture.rast, seq.dates)
# Applying the mask to plot only the amazon area.
soilm <- soilMoisture.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# define the zoom area
soilm.xy.zoom <- list(xmin=0.15e+06, xmax=0.2e+06, ymin=4.35e+06, ymax=4.5e+06, zoom=0.4)
# Plot
p.soilm.na <- myPlot(
  soilm, title = "Soil Moisture", 
  max_cell=1e7,
  x_angle=90,
  b_size=12,
  xy.zoom = soilm.xy.zoom
) + 
  scale_fill_hypso_c(
    name = TeX(r"($\textit{(mm)$})"),
    palette = "wiki-schwarzwald-cont", 
    na.value = "transparent")
p.soilm.na
```

## Plot the month of October 2020

```{r}
# Remove negative values.
soilm[soilm < 0] <- NA
# Plot
p.soilm <- myPlot(soilm, title = "Soil Moisture") + 
  scale_fill_hypso_c(
    name = TeX(r"($\textit{(mm)$})"),
    palette = "wiki-schwarzwald-cont", 
    na.value = "transparent")
p.soilm
```


### Zoom Plot after removing negative values

```{r}
# define the zoom area
soilm.xy.zoom <- list(xmin=0.15e+06, xmax=0.2e+06, ymin=4.35e+06, ymax=4.5e+06, zoom=0.4)
# Plot
p.soilm.na <- myPlot(
  soilm, title = "Soil Moisture", 
  max_cell=1e7,
  x_angle=90,
  b_size=12,
  na.color="black",
  xy.zoom = soilm.xy.zoom
) + 
  scale_fill_hypso_c(
    name = TeX(r"($\textit{(mm)$})"),
    palette = "wiki-schwarzwald-cont", 
    na.value = "transparent")
p.soilm.na
```

## Missing Data

```{r}
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl, cores=detectCores() - 1)

# Raster to datatable in parallel: one raster per thread
rasList <- foreach (
  ras_id=amaz.soilMoisture.list, 
  .packages=c('terra', 'sf', 'dplyr'), 
  .combine='c') %dopar% {
    # Read and rename raster
    ras <- rast(ras_id) %>% renameLayers(., 'soilmoisture_working_', '')
    # Replace negative value by `NA`
    ras[ras < 0] <- NA
    # Count the missing data
    ras.freq.na <- not.na(ras) %>%
      mask(amaz.basin.shp) %>%
      terra::freq(., digits=0, value=0, usenames=T)
  
    list(ras.freq.na)
  }
stopCluster(cl)

# Bind all per-raster into one dataframe
soilmoisture.freq.na <- rbindlist(rasList, fill=T, use.names=T)
# Order
colnames(soilmoisture.freq.na)[3] <- "soilmoisture_na"
soilmoisture.freq.na <- soilmoisture.freq.na[order(soilmoisture.freq.na$layer)]
soilmoisture.freq.na
```

#-

# Elevation data

## Import data

```{r}
# list of files
amaz.elevation.list <- list.files(
  paste0(path.data,"/5. Elevation/03. Working Data"),
  full.names=TRUE,
  pattern = ".tif$")
# Import data with "Terra"
elevation.rast <- rast(amaz.elevation.list)
elevation.rast
```

## Rename layer

```{r}
names(elevation.rast) <- "elevation"
elevation.rast
```

## Verification of the values

```{r}
# Verification of the values
elevation.minmax <- minmax(elevation.rast) %>% t() %>% as.data.frame()
elevation.minmax
```

## Plot the month of October 2020

```{r}
# Applying the mask to plot only the amazon area.
elev <- elevation.rast %>% mask(mask = amaz.basin.shp)
# Plot
wikicols <- hypsometric_tints_db %>% filter(pal == "wiki-2.0")
wikicols <- wikicols[wikicols$limit %between% c(-100, 6500),]
p.elev <- myPlot(elev, title = "Elevation") + 
  scale_fill_gradientn(
    name = TeX(r"($\textit{(m)}$)"),
    colors = wikicols$hex,
    values = scales::rescale(wikicols$limit),
    limit = range(wikicols$limit),
    na.value = "transparent")
p.elev
```

## Missing Data

```{r}
# Count the missing data
elevation.freq.na <- not.na(elevation.rast) %>% 
  mask(amaz.basin.shp) %>% 
  freq(., digits=0, value=0, usenames=T)
elevation.freq.na
```

#-

# LandSurfaceTemp data

## Import data

```{r}
# list of files
amaz.landSurfaceTemp.list <- list.files(
  paste0(path.data,"/6. LandSurfaceTemp/03. Working Data"),
  full.names=TRUE,
  pattern = ".tif$")
# Import data with "Terra"
landSurfaceTemp.rast <- rast(amaz.landSurfaceTemp.list)
landSurfaceTemp.rast
```

## Rename and order layers

```{r}
# Rename layers
landSurfaceTemp.rast <- renameLayers(landSurfaceTemp.rast, 'landsurftemp_working_', '')
# Order layers
landSurfaceTemp.rast <- landSurfaceTemp.rast[[ordered.names]]
landSurfaceTemp.rast
```

## Verification of the values

```{r}
# Verification of the values
landSurfaceTemp.minmax <- minmax(landSurfaceTemp.rast) %>% t() %>% as.data.frame()
landSurfaceTemp.minmax
```

## Plot the month of October 2020

```{r}
# Create a sequence date for 'rts' object
landSurfaceTemp.rts <- rts(landSurfaceTemp.rast, seq.dates)
# Applying the mask to plot only the amazon area.
lst <- landSurfaceTemp.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# Plot
p.lst <- myPlot(lst, title = "Land Surface Temperature") +
  scale_fill_whitebox_c(
    name = TeX(r"($\textit{(K)}$)"),
    palette = "muted", 
    na.value = "transparent")
p.lst
```

## Missing Data

```{r}
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl, cores=detectCores() - 1)

# Raster to datatable in parallel: one raster per thread
rasList <- foreach (
  ras_id=amaz.landSurfaceTemp.list, 
  .packages=c('terra', 'sf', 'dplyr'), 
  .combine='c') %dopar% {
    # Read and rename raster
    ras <- rast(ras_id) %>% renameLayers(., 'landsurftemp_working_', '')
    # Count the missing data
    ras.freq.na <- not.na(ras) %>%
      mask(amaz.basin.shp) %>%
      terra::freq(., digits=0, value=0, usenames=T)
  
    list(ras.freq.na)
  }
stopCluster(cl)

# Bind all per-raster into one dataframe
landsurftemp.freq.na <- rbindlist(rasList, fill=T, use.names=T)
# Order
colnames(landsurftemp.freq.na)[3] <- "landsurftemp_na"
landsurftemp.freq.na <- landsurftemp.freq.na[
  order(landsurftemp.freq.na$landsurftemp_na, decreasing=TRUE)]
landsurftemp.freq.na
```

### Plot 

```{r}
# Select the 4 months they have the most missing data
lst.4na <- landsurftemp.freq.na[1:4,]
lst.4na.month <- lst.4na$layer %>% paste0(., "_01") %>% gsub("_", "-", .)
lst.4na.rts <- subset(landSurfaceTemp.rts, lst.4na.month)
lst.4na.rast <- lst.4na.rts@raster %>% mask(mask = amaz.basin.shp) 
# Plot
p.lst.4na <- ggplot() +
  geom_spatvector(data = amaz.basin.shp$geometry, fill = "black", color = "gray40") +
  stat_spatraster(data = lst.4na.rast) +
  scale_x_continuous(labels = function(x) format(x, scientific = T, digits = 1)) +
  scale_y_continuous(labels = function(x) format(x, scientific = T, digits = 2)) + 
  ggtitle(label="Land Surface Temperature", subtitle=NULL) +
  coord_sf(datum = pull_crs(lst.4na.rast)) + 
  theme_bw(base_size=12) +
  scale_fill_whitebox_c(
    name = TeX(r"($\textit{(K)}$)"),
    palette = "muted", 
    na.value = "transparent") + 
  facet_wrap(~lyr, ncol = 2) +
  theme(axis.text.x = element_text(angle = 90))
p.lst.4na
```

#-

# Specific Humidity data

## Import data

```{r}
# list of files
amaz.humidity.list <- list.files(
  paste0(path.data,"/7. Specific Humidity/03. Working Data"),
  full.names=TRUE,
  pattern = ".tif$")
# Import data with "Terra"
humidity.rast <- rast(amaz.humidity.list)
humidity.rast
```

## Rename and order layers

```{r}
# Rename layers
humidity.rast <- renameLayers(humidity.rast, 'humidity_working_', '')
# Order layers
humidity.rast <- humidity.rast[[ordered.names]]
humidity.rast
```

## Verification of the values

```{r}
# Verification of the values
humidity.minmax <- minmax(humidity.rast) %>% t() %>% as.data.frame()
humidity.minmax
```

## Plot the month of October 2020

```{r}
# Create a sequence date for 'rts' object
humidity.rts <- rts(humidity.rast, seq.dates)
# Applying the mask to plot only the amazon area.
hum <- humidity.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# Plot
p.hum <- myPlot(hum, title = "Humidity") +
  scale_fill_cross_blended_c(
    name = TeX(r"($\textit{(kg_{water} / kg_{air})$})"),
    palette = "warm_humid", 
    na.value = "transparent")
p.hum
```

## Missing Data

```{r}
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl, cores=detectCores() - 1)

# Raster to datatable in parallel: one raster per thread
rasList <- foreach (
  ras_id=amaz.humidity.list, 
  .packages=c('terra', 'sf', 'dplyr'), 
  .combine='c') %dopar% {
    # Read and rename raster
    ras <- rast(ras_id) %>% renameLayers(., 'humidity_working_', '')
    # Count the missing data
    ras.freq.na <- not.na(ras) %>%
      mask(amaz.basin.shp) %>%
      terra::freq(., digits=0, value=0, usenames=T)
  
    list(ras.freq.na)
  }
stopCluster(cl)

# Bind all per-raster into one dataframe
humidity.freq.na <- rbindlist(rasList, fill=T, use.names=T)
# Order
colnames(humidity.freq.na)[3] <- "humidity_na"
humidity.freq.na <- humidity.freq.na[order(humidity.freq.na$layer)]
humidity.freq.na
```

### Zoom Plot

```{r}
# Applying the mask to plot only the amazon area.
hum <- humidity.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# define the zoom area
hum.xy.zoom <- list(xmin=1.e+06, xmax=1.4e+06, ymin=3.4e+06, ymax=4e+06, zoom=0.4)
# Plot
p.hum.na <- myPlot(
  hum, title = "Humidity", 
  max_cell=1e7,
  x_angle=90,
  b_size=12,
  na.color="black",
  xy.zoom = hum.xy.zoom
) +
  scale_fill_cross_blended_c(
    name = TeX(r"($\textit{(kg_{water} / kg_{air})$})"),
    palette = "warm_humid", 
    na.value = "transparent")
p.hum.na
```

#-

# Evapotranspiration data

## Import data

```{r}
# list of files
amaz.evapotranspiration.list <- list.files(
  paste0(path.data,"/8. Evapotranspiration/03. Working Data"),
  full.names=TRUE,
  pattern = ".tif$")
# Import data with "Terra"
evapotranspiration.rast <- rast(amaz.evapotranspiration.list)
evapotranspiration.rast
```

## Rename and order layers

```{r}
# Rename layers
evapotranspiration.rast <- renameLayers(
  evapotranspiration.rast, 'evapotranspiration_working_', '')
# Order layers
evapotranspiration.rast <- evapotranspiration.rast[[ordered.names]]
evapotranspiration.rast
```

## Verification of the values

```{r}
# Verification of the values
evapotranspiration.minmax <- minmax(evapotranspiration.rast) %>% 
  t() %>% 
  as.data.frame()
evapotranspiration.minmax
```

## Plot the month of October 2020

```{r}
# Create a sequence date for 'rts' object
evapotranspiration.rts <- rts(evapotranspiration.rast, seq.dates)
# Applying the mask to plot only the amazon area.
evapot <- evapotranspiration.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# Plot
p.evapot <- myPlot(evapot, title = "Evapotranspiration") +
  scale_fill_whitebox_c(
    name = TeX(r"($\textit{(kg/m^2s)}$)"),
    palette = "bl_yl_rd", 
    na.value = "transparent")
p.evapot
```

## Missing Data

```{r}
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl, cores=detectCores() - 1)

# Raster to datatable in parallel: one raster per thread
rasList <- foreach (
  ras_id=amaz.evapotranspiration.list, 
  .packages=c('terra', 'sf', 'dplyr'), 
  .combine='c') %dopar% {
    # Read and rename raster
    ras <- rast(ras_id) %>% renameLayers(., 'evapotranspiration_working_', '')
    # Count the missing data
    ras.freq.na <- not.na(ras) %>%
      mask(amaz.basin.shp) %>%
      terra::freq(., digits=0, value=0, usenames=T)
  
    list(ras.freq.na)
  }
stopCluster(cl)

# Bind all per-raster into one dataframe
evapotranspiration.freq.na <- rbindlist(rasList, fill=T, use.names=T)
# Order
colnames(evapotranspiration.freq.na)[3] <- "evapotranspiration_na"
evapotranspiration.freq.na <- 
  evapotranspiration.freq.na[order(evapotranspiration.freq.na$layer)]
evapotranspiration.freq.na
```

### Zoom Plot

```{r}
# Applying the mask to plot only the amazon area.
evapot <- evapotranspiration.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# define the zoom area
evapot.xy.zoom <- list(xmin=0.9e+06, xmax=1.4e+06, ymin=3e+06, ymax=3.6e+06, zoom=0.4)
# Plot
p.evapot.na <- myPlot(
  evapot, title = "Evapotranspiration", 
  max_cell=1e7,
  x_angle=90,
  b_size=12,
  na.color="black",
  xy.zoom = evapot.xy.zoom
) +
  scale_fill_whitebox_c(
    name = TeX(r"($\textit{(kg/m^2s)}$)"),
    palette = "bl_yl_rd", 
    na.value = "transparent")
p.evapot.na
```

#-

# Wind Speed data

## Import data

```{r}
# list of files
amaz.wind.list <- list.files(
  paste0(path.data,"/9. Wind Speed/03. Working Data"),
  full.names=TRUE,
  pattern = ".tif$")
# Import data with "Terra"
wind.rast <- rast(amaz.wind.list)
wind.rast
```

## Rename and order layers

```{r}
# Rename layers
wind.rast <- renameLayers(wind.rast, 'wind_working_', '')
# Order layers
wind.rast <- wind.rast[[ordered.names]]
wind.rast
```

## Verification of the values

```{r}
# Verification of the values
wind.minmax <- minmax(wind.rast) %>% t() %>% as.data.frame()
wind.minmax
```

## Plot the month of October 2020

```{r}
# Create a sequence date for 'rts' object
wind.rts <- rts(wind.rast, seq.dates)
# Applying the mask to plot only the amazon area.
wind <- wind.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# Plot
p.wind <- myPlot(wind, title = "Wind Speed") +
  scale_fill_hypso_c(
    name = TeX(r"($\textit{(m / s)}$)"),
    palette = "gmt_globe_bathy", 
    trans = "pseudo_log",
    direction = -1,
    na.value = "transparent")
p.wind
```

## Missing Data

```{r}
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl, cores=detectCores() - 1)

# Raster to datatable in parallel: one raster per thread
rasList <- foreach (
  ras_id=amaz.wind.list, 
  .packages=c('terra', 'sf', 'dplyr'), 
  .combine='c') %dopar% {
    # Read and rename raster
    ras <- rast(ras_id) %>% renameLayers(., 'wind_working_', '')
    # Count the missing data
    ras.freq.na <- not.na(ras) %>%
      mask(amaz.basin.shp) %>%
      terra::freq(., digits=0, value=0, usenames=T)
  
    list(ras.freq.na)
  }
stopCluster(cl)

# Bind all per-raster into one dataframe
wind.freq.na <- rbindlist(rasList, fill=T, use.names=T)
# Order
colnames(wind.freq.na)[3] <- "wind_na"
wind.freq.na <- wind.freq.na[order(wind.freq.na$layer)]
wind.freq.na
```

### Zoom Plot

```{r}
# Applying the mask to plot only the amazon area.
wind <- wind.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# define the zoom area
wind.xy.zoom <- list(xmin=0.9e+06, xmax=1.4e+06, ymin=3.4e+06, ymax=4e+06, zoom=0.4)
# Plot
p.wind.na <- myPlot(
  wind, title = "Wind Speed", 
  max_cell=1e7,
  x_angle=90,
  b_size=12,
  na.color="black",
  xy.zoom = wind.xy.zoom
) +
  scale_fill_hypso_c(
    name = TeX(r"($\textit{(m / s)}$)"),
    palette = "gmt_globe_bathy", 
    trans = "pseudo_log",
    direction = -1,
    na.value = "transparent")
p.wind.na
```

#-

# Air Temperature data

## Import data

```{r}
# list of files
amaz.airtemp.list <- list.files(
  paste0(path.data,"/10. Air Temperature/03. Working Data"),
  full.names=TRUE,
  pattern = ".tif$")
# Import data with "Terra"
airtemp.rast <- rast(amaz.airtemp.list)
airtemp.rast
```

## Rename and order layers

```{r}
# Rename layers
airtemp.rast <- renameLayers(airtemp.rast, 'airtemp_working_', '')
# Order layers
airtemp.rast <- airtemp.rast[[ordered.names]]
airtemp.rast
```

## Verification of the values

```{r}
# Verification of the values
airtemp.minmax <- minmax(airtemp.rast) %>% t() %>% as.data.frame()
airtemp.minmax
```

## Plot the month of October 2020

```{r}
# Create a sequence date for 'rts' object
airtemp.rts <- rts(airtemp.rast, seq.dates)
# Applying the mask to plot only the amazon area.
airtemp <- airtemp.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# Plot
p.airtemp <- myPlot(airtemp, title = "Air Temperature") +
  scale_fill_whitebox_c(
    name = TeX(r"($\textit{(K)}$)"),
    palette = "bl_yl_rd", 
    na.value = "transparent")
p.airtemp
```

## Missing Data

```{r}
cl <- makeCluster(detectCores() - 1)
registerDoParallel(cl, cores=detectCores() - 1)

# Raster to datatable in parallel: one raster per thread
rasList <- foreach (
  ras_id=amaz.airtemp.list, 
  .packages=c('terra', 'sf', 'dplyr'), 
  .combine='c') %dopar% {
    # Read and rename raster
    ras <- rast(ras_id) %>% renameLayers(., 'airtemp_working_', '')
    # Count the missing data
    ras.freq.na <- not.na(ras) %>%
      mask(amaz.basin.shp) %>%
      terra::freq(., digits=0, value=0, usenames=T)
  
    list(ras.freq.na)
  }
stopCluster(cl)

# Bind all per-raster into one dataframe
airtemp.freq.na <- rbindlist(rasList, fill=T, use.names=T)
# Order
colnames(airtemp.freq.na)[3] <- "airtemp_na"
airtemp.freq.na <- airtemp.freq.na[order(airtemp.freq.na$layer)]
airtemp.freq.na
```

### Zoom Plot

```{r}
# Applying the mask to plot only the amazon area.
airtemp <- airtemp.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
# define the zoom area
airtemp.xy.zoom <- list(xmin=0.9e+06, xmax=1.4e+06, ymin=3.4e+06, ymax=4e+06, zoom=0.4)
# Plot
p.airtemp.na <- myPlot(
  airtemp, title = "Air Temperature", 
  max_cell=1e7,
  x_angle=90,
  b_size=12,
  na.color="black",
  xy.zoom = airtemp.xy.zoom
) +
  scale_fill_whitebox_c(
    name = TeX(r"($\textit{(K)}$)"),
    palette = "bl_yl_rd", 
    na.value = "transparent")
p.airtemp.na
```

#-

# Merge all dataframes of missing values.

```{r}
# create the dataframe
amaz.na.df <- as.data.frame(ordered.names)
colnames(amaz.na.df) <- "layer"
# Merge the dataframes
amaz.na.df <- list(amaz.na.df, 
                   burntArea.freq.na[,-2], 
                   landCover.freq.na[,-2], 
                   precipitation.freq.na[,-2],
                   soilmoisture.freq.na[,-2],
                   elevation.freq.na[,-2],
                   landsurftemp.freq.na[,-2], 
                   humidity.freq.na[,-2], 
                   evapotranspiration.freq.na[,-2],
                   wind.freq.na[,-2], 
                   airtemp.freq.na[,-2]) %>% 
  reduce(full_join, by="layer")

amaz.na.df
```





