---
title: "Preparation_Amazon_data"
date: "2023-04-05"
output:
  html_document: 
    number_sections: yes
    fig_caption: yes
    toc: true
    toc_float: 
      collapsed: true
    theme: cerulean
    highlight: kate
    toc_depth: 5
    keep_md: yes
    df_print: paged
---

# Load libraries

```{r warning=FALSE, message=FALSE}
# load libraries
library(raster)
library(tictoc)
library(tidyverse)
library(terra)
library(sf)
library(RColorBrewer)
library(rts)
library(doParallel)
library(parallel)
library(foreach)
library(data.table)
library(tidyterra)
# Plot
library(RColorBrewer)
library(tidyterra)
library(scico)
library(wesanderson)
library(viridis)
library(scales)
library(latex2exp)
library(ggplot2)# Select a palette
library(ggpubr)
```

# Functions

```{r "Functions"}
# Function to save layers
saveLayers = function(dataRast, shapeFile, pathFile, strRemove,
                      x_min=xmin(dataRast), x_max=xmax(dataRast), 
                      y_min=ymin(dataRast), y_max=ymax(dataRast)){
  
  # Iterate all the layers
  for (i in 1:nlyr(dataRast)){
    print(i)
    dataRast.i <- dataRast[[i]]
    # Crop data
    dataRast.basin.i <- mask(dataRast.i, mask = shapeFile) %>%
      crop(ext(x_min, x_max, y_min, y_max))
    # Path of the file
    path <- gsub(strRemove, "", names(dataRast.basin.i)) %>%
      paste0(pathFile, ., ".tif")
    # Save file
    writeRaster(dataRast.basin.i, path, overwrite=TRUE)
    # Clear memory
    gc()
  }
}

# Function to rename layers
renameLayers <- function(dataRast, fileStart, prefix){
  # index of file name in the path
  id.date <- unlist(gregexpr(fileStart, sources(dataRast)[[1]])) + nchar(fileStart)
  # rename layers
  names(dataRast) <- 
    substr(sources(dataRast), id.date, (id.date+50)) %>%
    gsub(".tif", "_1", .) %>%
    as.Date("%Y_%m_%d") %>%
    format(., '%Y_%m') %>%
    paste0(prefix, .)
  
  return(dataRast)
}
#---- Plot ----
myPlot <- function(
  rast, 
  title=NULL, 
  sub_title=NULL, 
  theme=2, 
  na_value=NULL, 
  max_cell=1e6,
  x_angle=0,
  b_size=16,
  v_unit=c(2, 2, 2, 2) # c(t, r, b, l)
){
  # Prepare missing data
  if (!is.null(na_value)) {
    min.rast <- minmax(rast)[1]
    rast.na <- rast
    rast.na[is.na(rast.na)] <- min.rast - 1e3
    rast.na[rast.na >= min.rast] <- NA
    # levels(rast.na) <- data.frame(id=c(min.rast - 1e3), val=c('1'))
    rast.na <- rast.na %>% mask(mask = amaz.basin.shp)
    # rast.na[rast.na < min.rast] <- -1e-09
    p1 <- ggplot() +
      geom_spatraster(data = rast.na, maxcell = max_cell)
  }else{
    p1 <- ggplot()
  }
  p1 <- switch(
    theme,
    p1,
    p1 + theme_bw(base_size=b_size),
    p1 + theme_linedraw(base_size=b_size),
    p1 + theme_light(base_size=b_size),
    p1 + theme_minimal(base_size=b_size),
    p1 + theme_classic(base_size=b_size),
    p1 + theme_gray(base_size=b_size),
    p1 + theme_dark(base_size=b_size)
  )
  p1 <- p1  +
    geom_spatraster(data = rast, maxcell = max_cell) +
    geom_spatvector(data = amaz.basin.shp$geometry, fill = NA, color = "gray40") +
    scale_x_continuous(labels = function(x) format(x, scientific = T)) +
    scale_y_continuous(labels = function(x) format(x, scientific = T)) +
    ggtitle(label=title, subtitle=sub_title) +
    coord_sf(datum = pull_crs(rast)) +
    theme(
      axis.text.x = element_text(angle = x_angle)
      , plot.margin = unit(v_unit, "pt")
    )
  if (theme == 1) {p1 <- p1 + theme_void(base_size=b_size)}
  # coord_sf(crs = st_crs(4326))
  return(p1)
}
#--------------------------------------
myPlot2 <- function(rast, title=NULL, sub_title=NULL, theme=2, na_value="black"){
  # Load data
  title=NULL; sub_title=NULL; theme=2; na_value=-2
  rast <- landSurfaceTemp.rts[['2016-02-01']] %>% mask(mask = amaz.basin.shp)
  # Parasmeter
  max_cell <- 1e6
  # Prepare missing data
  if (!is.null(na.value)) {
    min.rast <- minmax(rast)[1]
    rast.na <- rast
    rast.na[is.na(rast.na)] <- min.rast - 1e3
    # rast.na[rast.na >= min.rast] <- NA
    # levels(rast.na) <- data.frame(id=c(min.rast - 1e3), val=c('1'))
    rast.na <- rast.na  %>% mask(mask = amaz.basin.shp)
    # rast.na[rast.na < min.rast] <- -2
    rast.na <- rast.na %>% mask(mask = amaz.basin.shp)
    # levels(rast.na) <- data.frame(id=c(min.rast - 1e03), val='1')
    p1 <-  ggplot()  +
      geom_spatraster(data = rast.na, maxcell = max_cell) +
      scale_fill_gradientn(colours = na_value, na.value = NA) +
      # scale_fill_manual(values = na_value) +
      theme(
        legend.position = "none",
        axis.text        = element_blank(),
        axis.ticks       = element_blank(),
        axis.title       = element_blank(),
        panel.background = element_blank())
  }
  p2 <- ggplot()  +
    geom_spatraster(data = rast, maxcell = max_cell) +
    geom_spatvector(data = amaz.basin.shp$geometry, fill = NA, color = "gray40") +
    scale_x_continuous(labels = function(x) format(x, scientific = T)) +
    scale_y_continuous(labels = function(x) format(x, scientific = T)) +
    ggtitle(label=title, subtitle=sub_title) +
    coord_sf(datum = pull_crs(rast))
    # coord_sf(crs = st_crs(4326))
  p2 <- switch(
    theme,
    p2 + theme_void(),
    p2 + theme_bw(),
    p2 + theme_linedraw(),
    p2 + theme_light(),
    p2 + theme_minimal(),
    p2 + theme_classic(),
    p2 + theme_gray(),
    p2 + theme_dark()
  )
  return(p2)
  
  my.colors <- c("mediumblue", "mediumseagreen", "firebrick")

  ggplot()  +
    geom_spatraster(data = rast.na, maxcell = max_cell) +
    geom_spatvector(data = amaz.basin.shp$geometry, fill = NA, color = "gray40") +
    geom_spatraster(data = rast, maxcell = max_cell) +
    scale_x_continuous(labels = function(x) format(x, scientific = T)) +
    scale_y_continuous(labels = function(x) format(x, scientific = T)) +
    ggtitle(label=title, subtitle=sub_title) +
    coord_sf(datum = pull_crs(rast)) + 
    theme_bw() +
    # scale_fill_manual(name = "Burnt Area", values = my.colors, na.translate=FALSE)
    
    
    
    p2 + scale_fill_whitebox_c(
      name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
      palette = "muted", 
      na.value = "transparent"
    )
  
  
  
}
```

# Test

```{r}
lst <- landSurfaceTemp.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
rast <- lst
myPlot(lst) +
  scale_fill_whitebox_c(
    name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    palette = "muted", 
    na.value = "transparent")
```





# Initialization

```{r}
my.path <- "/Users/ABIDM/Documents/Project/Code_official_data/"

# Import shape file
amaz.basin.shp <- st_read("/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/0. Amazon_shapefile/projected/amazon_shp_projected.shp")

# path of variables
burntArea.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/1. Burnt Area/03. Working Data"
landCover.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/2. Land Cover/03. Working Data"
precip.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/3. Precipitation/03. Working Data"
soilMoisture.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/4. Soil Moisture/03. Working Data"
elevation.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/5. Elevation/03. Working Data"
LandSurfaceTemp.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/6. LandSurfaceTemp/03. Working Data"
humidity.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/7. Specific Humidity/03. Working Data"
evapotranspiration.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/8. Evapotranspiration/03. Working Data"
wind.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/9. Wind Speed/03. Working Data"
airtemp.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/10. Air Temperature/03. Working Data"

# Inverse basin shape


x_min=-1.25e+06; x_max=0.05e+06; y_min=st_bbox(amaz.basin.shp)$ymin; y_max=2.4e+06    # South of Amazon

# Number of cores to use
# nbrCores <- detectCores() - 1
nbrCores <- 45

# Color palette
pal <- colorRampPalette(c("mediumblue", "forestgreen", "firebrick"))
pals <- unique(hypsometric_tints_db$pal)
```

#=====================

# Import data

```{r "1.Import data"}
#---- Burnt Area data ----
burntArea.rast <- 
  list.files(burntArea.path, full.names=TRUE, pattern = ".tif$") %>%
  rast()
#---- Land Cover ----
landCover.rast <- 
  list.files(landCover.path, full.names=TRUE, pattern = ".tif$") %>%
  rast()
#---- Precipitation ----
precipitation.rast <- 
  list.files(precip.path, full.names=TRUE, pattern = ".tif$") %>% 
  rast()
#----- Soil Moisture ----
soilMoisture.rast <- 
  list.files(soilMoisture.path, full.names=TRUE, pattern = ".tif$") %>%
  rast()
#---- Elevation ----
elevation.rast <- 
  list.files(elevation.path,  full.names=TRUE, pattern = ".tif$") %>%
  rast()
#---- LandSurfaceTemp ----
landSurfaceTemp.rast <- 
  list.files(LandSurfaceTemp.path, full.names=TRUE, pattern = ".tif$") %>% 
  rast()
#---- Specific Humidity ----
humidity.rast <- 
  list.files(humidity.path, full.names=TRUE, pattern = ".tif$") %>% 
  rast()
#---- Evapotranspiration ----
evapotranspiration.rast <- 
  list.files(evapotranspiration.path, full.names=TRUE, pattern = ".tif$") %>%
  rast()
#---- Wind Speed ----
wind.rast <- 
  list.files(wind.path, full.names=TRUE, pattern = ".tif$") %>%
  rast()
#---- Air Temperature ----
airtemp.rast <- 
  list.files(airtemp.path, full.names=TRUE, pattern = ".tif$") %>%
  rast()
```

# Modifications

```{r "2.Modifications"}
#----- Soil Moisture ----
# # Repalce negative values by `NA` for all data
# soilMoisture.rast[soilMoisture.rast < 0] <- NA
# # Save/Load raster file
# writeRaster(soilMoisture.rast, "Amazon_new_data/4. Soil Moisture/soilmoisture_working_na.tif", overwrite=TRUE)
soilMoisture.rast <- rast("Amazon_new_data/4. Soil Moisture/soilmoisture_working_na.tif")
# Remove September and November 2012
soilMoisture238.rast <- soilMoisture.rast[[!(names(soilMoisture.rast) %in% c("2012_07", "2012_09"))]]
# # Save/Load raster file
# writeRaster(soilMoisture238.rast, "Amazon_new_data/4. Soil Moisture/soilmoisture_working_238.tif", overwrite=TRUE)
soilMoisture.rast <- rast("Amazon_new_data/4. Soil Moisture/soilmoisture_working_238.tif")
```

# Rename and order layers

```{r}
ordered.names<- 
  seq(as.Date("2001-1-1"), as.Date("2020-12-1"), by = "month") %>% 
  format(., '%Y_%m') %>% 
  setdiff(., c("2012_07", "2012_09"))
#---- Burnt Area data ----
burntArea.rast <- renameLayers(burntArea.rast, 'burntarea_working_', '')
burntArea.rast <- burntArea.rast[[ordered.names]]
#---- Land Cover ----
landCover.rast <- renameLayers(landCover.rast, 'landcover_working_', '')
landCover.rast <- landCover.rast[[ordered.names]]
#---- Precipitation ----
precipitation.rast <- renameLayers(precipitation.rast, 'precipitation_working_', '')
precipitation.rast <- precipitation.rast[[ordered.names]]
#----- Soil Moisture ----
# soilMoisture.rast <- renameLayers(soilMoisture.rast, 'soilmoisture_working_', '')
# names(soilMoisture.rast) <- ordered.names
#---- Elevation ----
names(elevation.rast) <- "elevation"
#---- LandSurfaceTemp ----
landSurfaceTemp.rast <- renameLayers(landSurfaceTemp.rast, 'landsurftemp_working_', '')
landSurfaceTemp.rast <- landSurfaceTemp.rast[[ordered.names]]
#---- Specific Humidity ----
humidity.rast <- renameLayers(humidity.rast, 'humidity_working_', '')
humidity.rast <- humidity.rast[[ordered.names]]
#---- Evapotranspiration ----
evapotranspiration.rast <- renameLayers(evapotranspiration.rast, 'evapotranspiration_working_', '')
evapotranspiration.rast <- evapotranspiration.rast[[ordered.names]]
#---- Wind Speed ----
wind.rast <- renameLayers(wind.rast, 'wind_working_', '')
wind.rast <- wind.rast[[ordered.names]]
#---- Air Temperature ----
airtemp.rast <- renameLayers(airtemp.rast, 'airtemp_working_', '')
airtemp.rast <- airtemp.rast[[ordered.names]]
```

# Create `rts` object

```{r}
# Dates
seq.dates <- seq(as.Date("2001-1-1"), as.Date("2020-12-1"), by = "month")
seq.dates <- seq.dates[!((seq.dates == "2012-07-01") | (seq.dates == "2012-09-01"))]
date.2020.12 <- seq(as.Date("2020-12-1"), as.Date("2020-12-1"), by = "month")
# Create 'rts' objects
burntArea.rts <- rts(burntArea.rast, seq.dates)
landCover.rts <- rts(landCover.rast, seq.dates)
precipitation.rts <- rts(precipitation.rast, seq.dates)
soilMoisture.rts <- rts(soilMoisture.rast, seq.dates)
elevation.rts <- rts(elevation.rast, date.2020.12)
landSurfaceTemp.rts <- rts(landSurfaceTemp.rast, seq.dates)
humidity.rts <- rts(humidity.rast, seq.dates)
evapotranspiration.rts <- rts(evapotranspiration.rast, seq.dates)
wind.rts <- rts(wind.rast, seq.dates)
airtemp.rts <- rts(airtemp.rast, seq.dates)
```

# Plot

## Burnt Area

```{r "1.1.Burnt Area", warning=FALSE}
ba <- burntArea.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
levels(ba) <- data.frame(id=c(-2, 0, 1), val=c('-2', '0', '1'))
my.colors <- c("mediumblue", "mediumseagreen", "firebrick")
p.ba <- myPlot(ba, title = "Burnt Area", v_unit=c(3, 55, 3, 0)) +
  scale_fill_manual(
    name = NULL, 
    values = my.colors, 
    na.translate=FALSE
  ) 
p.ba
```

```{r "1.2.Q1", message=FALSE}
Q1 <- rast("/Users/ABIDM/Documents/Project/Code_official_data/Amazon_new_data/1. Burnt Area/ras/Q1_238.tif") 
levels(Q1) <- data.frame(id=c(-2, 0, 1), val=c('-2', '0', '1'))
my.colors <- c("mediumblue", "mediumseagreen", "firebrick")
p.q1 <- myPlot(Q1, "Maximum value for each cell", theme=1) +
  scale_fill_manual(name = "Burnt Area", values = my.colors, na.translate=FALSE)
p.q1
```
###Burnt Area in October 2020

```{r}
ba.2020.10 <- burntArea.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
levels(ba.2020.10) <- data.frame(id=c(-2, 0, 1), val=c('-2', '0', '1'))
my.colors <- c("mediumblue", "mediumseagreen", "firebrick")
 myPlot(ba.2020.10, "Burnt Area in October 2020") +
  scale_fill_manual(name = "Burnt Area", values = my.colors, na.translate=FALSE)
 
 
ba.2020.10.dt <- as.data.table(ba.2020.10, cell=T, xy=T)

ggplot(ba.2020.10.dt, aes(x=val))+
  geom_bar(fill="lightpink1") +
  labs(x="Burnt Area in October 2020") +
  geom_text(aes(label = ..count..), stat = "count")+
  theme_bw()
#---
ggplot(data = ba.2020.10.dt, aes(x = val, fill = val)) +
  geom_text(aes(label = ..count..), stat = "count") +
  scale_fill_manual(values = my.colors) +
  geom_bar() + 
  theme_bw()


  
  
 freq(ba.2020.10) 
 
 
```

```{r}
ggplot(data = ba.2020.10.dt, aes(x = val)) + 
  geom_bar(stat = "count", 
           aes(fill = val), position = "dodge") + 
  labs(title="Burnt Area in October 2020", x="Burnt Area") +
  coord_flip() + 
  stat_count(geom = "text", 
             aes(label = ..count..),
             position=position_stack(vjust=0.5),
             colour = "black", size = 3.5) + 
  theme_bw()
```

```{r}
my.colors <- c("mediumblue", "mediumseagreen", "firebrick")
ggplot(data = ba.2020.10.dt, aes(x = val)) + 
  geom_bar(stat = "count", 
           aes(fill = val), position = "dodge") + 
  labs(title="Burnt Area in October 2020", x="Burnt Area") +
  coord_flip() + 
  scale_fill_manual(name = "Burnt Area", values = my.colors) +
  stat_count(geom = "text", 
             aes(label = ..count..),
             position=position_stack(vjust=0.5),
             colour = "black", size = 3.5) + 
  theme_bw()
```


## Land Cover

```{r "2.1.Land Cover"}
lc <- landCover.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)  
levels(lc) <- data.frame(id=0:10, val=c('0', '1', '2', '3', '4', '5', '6', '7', '8', '9', '10'))
p.lc <- myPlot(lc, title = "Land Cover", v_unit=c(3, 55, 3, 0)) + 
  scale_fill_hypso_d(
    name = NULL, 
    palette = "colombia_hypso", 
    na.translate=FALSE)
p.lc
```

## Precipitation

```{r "3.1.Precipitation"}
prec <- precipitation.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
p.prec <- myPlot(prec, title = "Precipitation", v_unit=c(3, 43, 3, 0)) + 
  scale_fill_scico(
    # name = TeX(r"(\overset{Precipitation}{$\textit{(mm/hr)}$})"),
    name = TeX(r"($\textit{(mm/hr)}$)"),
    palette = "lapaz", 
    direction = -1,
    trans = "pseudo_log",
    breaks = c(0,10,50,200,550),
    na.value = "transparent")
p.prec
```

## Soil Moisture

```{r "4.1.Soil Moisture"}
soilm <- soilMoisture.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
p.soilm <- myPlot(soilm, title = "Soil Moisture", v_unit=c(3, 50, 3, 0)) + 
  scale_fill_hypso_c(
    # name = TeX(r"(\overset{Soil Moisture }{$\textit{(mm)}$})"),
    name = TeX(r"($\textit{(mm)$})"),
    palette = "wiki-schwarzwald-cont", 
    na.value = "transparent")
p.soilm
```

```{r}
myPlot(soilm, max_cell=1e7) + 
  scale_fill_hypso_c(
    name = TeX(r"(\overset{Soil Moisture }{$\textit{(mm)}$})"),
    palette = "wiki-schwarzwald-cont", 
    na.value = "transparent")
```

## Elevation

```{r "5.1.Elevation"}
elev <- elevation.rts[['2020-12-01']] %>% mask(mask = amaz.basin.shp)
wikicols <- hypsometric_tints_db %>% filter(pal == "wiki-2.0")
wikicols <- wikicols[wikicols$limit %between% c(-100, 6500),]
p.elev <- myPlot(elev, title = "Elevation", v_unit=c(3, 45, 3, 0)) + 
  scale_fill_gradientn(
    # name = TeX(r"(\overset{Elevation }{$\textit{(m)}$})"),
    name = TeX(r"($\textit{(m)}$)"),
    colors = wikicols$hex,
    values = scales::rescale(wikicols$limit),
    limit = range(wikicols$limit),
    na.value = "transparent")
p.elev
```

## LandSurfaceTemp

```{r "6.1.LandSurfaceTemp"}
lst <- landSurfaceTemp.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
p.lst <- myPlot(lst, title = "Land Surface Temperature", v_unit=c(3, 40, 3, 0)) +
  scale_fill_whitebox_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{(K)}$)"),
    palette = "muted", 
    na.value = "transparent")
p.lst
```

### Plot missing data sample

```{r "6.1.LandSurfaceTemp"}
lst2 <- landSurfaceTemp.rts[['2002-02-01']] %>% mask(mask = amaz.basin.shp)
p.lst2 <- myPlot(lst, na_value = "black", title = "Land Surface Temperature", v_unit=c(3, 40, 3, 0)) +
  scale_fill_whitebox_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{(K)}$)"),
    palette = "muted", 
    na.value = "transparent")
p.lst2
```

## Specific Humidity

```{r "7.1.Specific Humidity"}
hum <- humidity.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
p.hum <- myPlot(hum, title = "Humidity", v_unit=c(3, 17, 3, 0)) +
  scale_fill_cross_blended_c(
    # name = TeX(r"(\overset{Humidity }{$\textit{(kg_{water} / kg_{air})}$})"),
    name = TeX(r"($\textit{(kg_{water} / kg_{air})$})"),
    palette = "warm_humid", 
    na.value = "transparent")
p.hum
```

## Evapotranspiration

```{r "8.1.Evapotranspiration"}
evapot <- evapotranspiration.rts[['2020-01-01']] %>% mask(mask = amaz.basin.shp)
p.evapot <- myPlot(evapot, title = "Evapotranspiration", v_unit=c(3, 40, 3, 0)) +
  scale_fill_whitebox_c(
    # name = TeX(r"(\overset{Evapotranspiration}{$\textit{(kg/m^2s)}$})"),
    name = TeX(r"($\textit{(kg/m^2s)}$)"),
    palette = "bl_yl_rd", 
    na.value = "transparent")
p.evapot
```

## Wind Speed

```{r "9.1.Wind Speed"}
wind <- wind.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
p.wind <- myPlot(wind, title = "Wind Speed", v_unit=c(3, 57, 3, 0)) +
  scale_fill_hypso_c(
    # name = TeX(r"(\overset{Wind Speed }{$\textit{(m / s)}$})"),
    name = TeX(r"($\textit{(m / s)}$)"),
    palette = "gmt_globe_bathy", 
    trans = "pseudo_log",
    direction = -1,
    na.value = "transparent")
p.wind
```

## Air Temperature

```{r "10.1.Air Temperature"}
airtemp <- airtemp.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
p.airtemp <- myPlot(airtemp, title = "Air Temperature", v_unit=c(3, 50, 3, 0)) +
  scale_fill_whitebox_c(
    name = TeX(r"($\textit{(K)}$)"),
    palette = "bl_yl_rd", 
    na.value = "transparent")
p.airtemp
```

# Create and plot zones

```{r, "11.Zones"}
v <- ext(amaz.basin.shp)
x.min <- v$xmin[[1]] - 1; x.max <- v$xmax[[1]] + 1; 
y.min <- v$ymin[[1]] - 1; y.max <- v$ymax[[1]] + 1; 

XY <- list(
  c(x.min, -0.84e+06, y.min, y.max),                       # z1
  c(-0.84e+06, +0.55e+06, 3.75e+06, y.max),                # z2
  c(-0.84e+06, +0.55e+06, 2.6e+06, 3.75e+06),              # z3
  c(-0.84e+06, -0.05e+06, 2.22e+06, 2.6e+06),              # z4
  c(-0.84e+06, -0.51e+06, y.min, 2.22e+06),                # z5
  c(-0.51e+06, -0.05e+06, y.min, 2.22e+06),                # z6
  c(-0.05e+06, 0.6e+06, y.min, 2.6e+06),                   # z7
  c(0.6e+06, 0.95e+06, y.min, 2.6e+06),                    # z8
  c(+0.55e+06, x.max, 3.24e+06, y.max),                    # z9
  c(+0.55e+06, x.max, 2.86e+06, 3.24e+06),                 # z10
  c(+0.55e+06, 0.95e+06, x.max, y.min, 2.6e+06, 2.86e+06)  # z11
)

nz <- length(XY)
zl <- list()
for (i in 1:(nz-1)){
  zl[[i]] <- rbind(
    c(XY[[i]][1], XY[[i]][3]), 
    c(XY[[i]][2], XY[[i]][3]), 
    c(XY[[i]][2], XY[[i]][4]), 
    c(XY[[i]][1], XY[[i]][4])
  ) %>% 
    cbind(object=i, part=1, ., hole=0) %>% 
    as.data.table()
}

i <- nz
zl[[i]] <- rbind(
  c(XY[[i]][2], XY[[i]][4]), 
  c(XY[[i]][3], XY[[i]][4]), 
  c(XY[[i]][3], XY[[i]][6]), 
  c(XY[[i]][1], XY[[i]][6]), 
  c(XY[[i]][1], XY[[i]][5]), 
  c(XY[[i]][2], XY[[i]][5])
) %>% 
  cbind(object=i, part=1, ., hole=0) %>% 
  as.data.table()

z <- rbindlist(zl, use.names=T) %>% as.matrix()
colnames(z)[3:4] <- c('x', 'y')

zones.shp <- vect(z, "polygons")
crs(zones.shp) <- crs(amaz.basin.shp)
#---- Save ----
writeVector(p, "Amazon_new_data/1. Burnt Area/Zones_shp/Zones.shp", overwrite=F)
zones.shp <- st_read("Amazon_new_data/1. Burnt Area/Zones_shp/Zones.shp")

#---- Plot ----
my.labels <-  c(
  TeX(r"($\textit{Z_1}$)")
  , TeX(r"($\textit{Z_2}$)")
  , TeX(r"($\textit{Z_3}$)")
  , TeX(r"($\textit{Z_4}$)")
  , TeX(r"($\textit{Z_5}$)")
  , TeX(r"($\textit{Z_6}$)")
  , TeX(r"($\textit{Z_7}$)")
  , TeX(r"($\textit{Z_8}$)")
  , TeX(r"($\textit{Z_9}$)")
  , TeX(r"($\textit{Z_{10}}$)")
  , TeX(r"($\textit{Z_{11}}$)")
)
x.coor <- c(-1.50e+06, -0.50e+06, -0.50e+06, -0.5e+06, -0.70e+06, -0.20e+06, 0.30e+06,  0.8e+06, 1.30e+06, 1.60e+06, 1.35e+06)
y.coor <- c( 3.05e+06,  4.05e+06,  3.05e+06,  2.4e+06,  1.85e+06,  1.85e+06, 1.90e+06, 2.00e+06, 3.75e+06, 3.00e+06, 2.50e+06)

p.zones <- myPlot(Q1, "Maximum value for each cell", theme=1) +
  geom_spatvector(data = zones.shp, fill = NA, color = "black") +
  coord_sf(datum = pull_crs(Q1)) + 
  scale_fill_manual(name = "Burnt Area", values = alpha(my.colors, 0.75), na.translate=FALSE) +
  annotate("text", label = my.labels, x = x.coor, y = y.coor, size = 6, colour = "black")
p.zones
```

# Missing data

```{r}
load("/Users/ABIDM/Documents/Project/Code_official_data/R_data/amaz_na_df.Rdata")
amaz.na.df.12lst <- amaz.na.df[order(amaz.na.df$landsurftemp_na, decreasing=TRUE),][1:4,]
month.na.12lst <- amaz.na.df.12lst$layer %>% paste0(., "_01") %>% gsub("_", "-", .)
rast <- subset(landSurfaceTemp.rts, month.na.12lst)
lst.max12na.rast <- rast@raster %>% mask(mask = amaz.basin.shp) 
# Save / Load raster file
# writeRaster(lst.max12na.rast, "Amazon_new_data/6. LandSurfaceTemp/landsurftemp_na_max12months.tif", overwrite=F)
# lst.max12na.rast <- rast("Amazon_new_data/6. LandSurfaceTemp/landsurftemp_na_max12months.tif")

# Plot
p.lst.max12na <- ggplot() +
  stat_spatraster(data = lst.max12na.rast) +
  geom_spatvector(data = amaz.basin.shp$geometry, fill = NA, color = "gray40") +
  scale_x_continuous(labels = function(x) format(x, scientific = T, digits = 1)) +
  scale_y_continuous(labels = function(x) format(x, scientific = T, digits = 2)) + 
  ggtitle(label="Land Surface Temperature", subtitle=NULL) +
  coord_sf(datum = pull_crs(lst.max12na.rast)) + 
  theme_bw(base_size=10) +
  scale_fill_whitebox_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{(K)}$)"),
    palette = "muted", 
    na.value = "transparent") + 
  facet_wrap(~lyr, ncol = 2) +
  theme(axis.text.x = element_text(angle = 90))
p.lst.max12na
```

# Data preparation

```{r}
Q0 <- rast(paste0(my.path,"Amazon_new_data/1. Burnt Area/ras/Q0_238.tif"))
Q1 <- rast(paste0(my.path,"Amazon_new_data/1. Burnt Area/ras/Q1_238.tif"))

# Plot
levels(Q0) <- data.frame(id=c(-2, 0, 1), cover=c('-2', '0', '1'))
levels(Q1) <- data.frame(id=c(-2, 0, 1), cover=c('-2', '0', '1'))
my.colors <- c("mediumblue", "mediumseagreen", "firebrick")

myPlot(
  Q0
  , "Plot of the minimum value for each cell"
  # , paste0("Total number of cells = ", length( cells(Q0) )) 
  , theme = 1
) +
  scale_fill_manual(name = "Burnt Area", values = my.colors, na.translate=FALSE)

myPlot(
  Q1
  , "Plot of the maximum value for each cell"
  # , paste0("Total number of cells = ", length( cells(Q1) )) 
  , theme = 1
) +
  scale_fill_manual(name = "Burnt Area", values = my.colors, na.translate=FALSE)
```

## Split data into 3 parts

```{r, message=FALSE}
ras0 <- ras1 <- ras2 <- Q1

# Create a pattern for burntArea = 0
ras0[ras0 %in% c(-2, 1)] <- NA
ras0[ras0 == 0] <- 1


# Create a pattern for burntArea = 1
ras1[ras1 %in% c(-2, 0)] <- NA

# Create a pattern for burntArea = -2
ras2[ras2 %in% c(0, 1)] <- NA
ras2[ras2 == -2] <- 1

# Plot
levels(ras2) <- data.frame(id=1, cover=c('-2'))
levels(ras0) <- data.frame(id=1, cover=c('0'))
levels(ras1) <- data.frame(id=1, cover=c('1'))

myPlot(
  ras2
  , "Cells always corresponding to 'water' regions"
  , paste0("Total number of cells = ", 
           formatC(length(cells(ras2)), format="f", big.mark=",", digits=0)  )
  , theme = 1
) +
  scale_fill_manual(name = "Burnt Area", values = my.colors[1], na.translate=FALSE)

myPlot(
  ras0
  , "Cells with 'no fire' events"
  , paste0("Total number of cells = ", 
           formatC(length(cells(ras0)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
  scale_fill_manual(name = "Burnt Area", values = my.colors[2], na.translate=FALSE)

myPlot(
  ras1
  , "Cells with at least a 'fire' event"
  , paste0("Total number of cells = ", 
           formatC(length(cells(ras1)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
  scale_fill_manual(name = "Burnt Area", values = my.colors[3], na.translate=FALSE)
```

```{r}
# Load 
ras1 <- rast(paste0(my.path,"Amazon_new_data/1. Burnt Area/ras/ras1_fires.tif"))
ras2 <- rast(paste0(my.path,"Amazon_new_data/1. Burnt Area/ras/ras2_alwaysWater.tif"))
ras2.0 <- rast(paste0(my.path,"Amazon_new_data/1. Burnt Area/ras/ras2_0_ChangeWaterNofire.tif"))
ras0.0 <- rast(paste0(my.path,"Amazon_new_data/1. Burnt Area/ras/ras0_0_alwaysNoFire.tif"))

# Plot
levels(ras2) <- data.frame(id=1, cover=c('-2'))
levels(ras2.0) <- data.frame(id=1, cover=c('-2'))
levels(ras0.0) <- data.frame(id=1, cover=c('0'))
levels(ras1) <- data.frame(id=1, cover=c('1'))

myPlot(
  ras2
  , "Cells always corresponding to 'water' regions"
  , paste0("Total number of cells = ", 
           formatC(length(cells(ras2)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
  scale_fill_manual(name = "Burnt Area", values = my.colors[1], na.translate=FALSE)

myPlot(
  ras2.0
  , "Cells may correspond to  'water' or 'no fire' regions"
  , paste0("Total number of cells = ", 
           formatC(length(cells(ras2.0)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
  scale_fill_manual(name = "Burnt Area", values = my.colors[1], na.translate=FALSE)

myPlot(
  ras0.0
  , "Cells always corresponding to 'no fire' regions"
  , paste0("Total number of cells = ", 
           formatC(length(cells(ras0.0)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
  scale_fill_manual(name = "Burnt Area", values = my.colors[2], na.translate=FALSE)

myPlot(
  ras1
  , "Cells with at least a 'fire' event"
  , paste0("Total number of cells = ", 
           formatC(length(cells(ras1)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
  scale_fill_manual(name = "Burnt Area", values = my.colors[3], na.translate=FALSE)
```

```{r}
landSurfaceTemp0.rand <-rast(paste0(my.path,"Amazon_new_data/6. LandSurfaceTemp/ras/landSurfaceTemp0_rand.tif"))
landSurfaceTemp0.rand.NA.mask <-rast(paste0(my.path,"Amazon_new_data/6. LandSurfaceTemp/ras/landSurfaceTemp0_rand_NA_mask.tif"))

rand.rast0 <- ras0.0
values(rand.rast0) <- sample(1:238, ncell(rand.rast0), replace=TRUE)
rand.rast0 <- selectRange(rand.rast0, ras0.0)
names(rand.rast0) <- "Month_nbr"

temp <- landSurfaceTemp0.rand
temp[!is.na(temp)] <- 1
landSurfaceTemp0.ind.rand <- selectRange(rand.rast0, temp)

myPlot(
  landSurfaceTemp0.rand
  , "Random values from Land Surface Temperature"
  , paste0("Total number of cells = ", 
           formatC(length(cells(landSurfaceTemp0.rand)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
scale_fill_hypso_c(
  name = TeX(r"(Number of month)"),
  palette = "wiki-schwarzwald-cont", 
  na.value = "transparent")
  
myPlot(
  landSurfaceTemp0.rand.NA.mask
  , "Missing data in 'Land Surface Temprature'"
  , paste0("Total number of cells = ", 
           formatC(length(cells(landSurfaceTemp0.rand.NA.mask)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
scale_fill_hypso_c(
  name = TeX(r"(Number of month)"),
  palette = "wiki-schwarzwald-cont", 
  na.value = "transparent")

myPlot(
  landSurfaceTemp0.ind.rand
  , "Pattern for cells with 'no fire' and no missing data"
  , paste0("Total number of cells = ", 
           formatC(length(cells(landSurfaceTemp0.ind.rand)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
scale_fill_hypso_c(
  name = TeX(r"(Number of month)"),
  palette = "wiki-schwarzwald-cont", 
  na.value = "transparent")
```

```{r}
load(paste0(my.path,"Amazon_new_data/6. LandSurfaceTemp/05. cell_0/landSurfaceTemp0_NA_sample_dt.Rdata"))
landSurfaceTemp0.rand.NA.mask[is.na(landSurfaceTemp0.rand.NA.mask)] <- -1
landSurfaceTemp0.allcell.NA.dt <- terra::as.data.frame(landSurfaceTemp0.rand.NA.mask, xy=T, cells=T)
landSurfaceTemp0.allcell.NA.dt$x <- floor(landSurfaceTemp0.allcell.NA.dt$x)
landSurfaceTemp0.allcell.NA.dt$y <- floor(landSurfaceTemp0.allcell.NA.dt$y)
landSurfaceTemp0.allcell.NA.dt <- as.data.table(landSurfaceTemp0.allcell.NA.dt)

landSurfaceTemp0.allcell.NA.sample.dt <- merge(x=landSurfaceTemp0.allcell.NA.dt[, c('cell', 'x', 'y')], y=landSurfaceTemp0.NA.sample.dt[, c('cell', 'x', 'y', 'Month_nbr')], by = c('cell', 'x', 'y'), all = TRUE)
landSurfaceTemp0.allcell.NA.sample.dt

values(landSurfaceTemp0.rand.NA.mask) <- landSurfaceTemp0.allcell.NA.sample.dt$Month_nbr

myPlot(
  landSurfaceTemp0.rand.NA.mask
  , "Pattern for cells with 'no fire' and missing data"
  , paste0("Total number of cells = ", 
           formatC(length(cells(landSurfaceTemp0.rand.NA.mask)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
scale_fill_hypso_c(
  name = TeX(r"(Number of month)"),
  palette = "wiki-schwarzwald-cont", 
  na.value = "transparent")
```

```{r}
lst <- landSurfaceTemp.rts[['2020-10-01']] %>% mask(mask = amaz.basin.shp)
myPlot(
  lst
  , title = "Land Surface Temperature"
  , v_unit=c(3, 40, 3, 0)
  , theme = 1
) +
  scale_fill_whitebox_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{(K)}$)"),
    palette = "muted", 
    na.value = "transparent")
```

```{r}
rand.p2.all.rast <- rast(paste0(my.path,"Amazon_new_data/1. Burnt Area/ras_pattern/rand_p2_all_rast.tif"))

myPlot(
  rand.p2.all.rast
  , title = "Pattern for cells that may correspond to 'water' or \n'no fire' regions"
  , paste0("Total number of cells = ", 
           formatC(length(cells(rand.p2.all.rast)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
  scale_fill_hypso_c(
    name = TeX(r"(Number of month)"),
    palette = "wiki-schwarzwald-cont", 
    na.value = "transparent")
```

```{r}
rand.p0.p2.all.rast <- rast(paste0(my.path,"Amazon_new_data/1. Burnt Area/ras_pattern/rand_p0_p2_all_rast.tif"))

myPlot(
  rand.p0.p2.all.rast
  , title = "Pattern for cells that didn't have any fire event"
  , paste0("Total number of cells = ", 
           formatC(length(cells(rand.p0.p2.all.rast)), format="f", big.mark=",", digits=0) )
  , theme = 1
) +
  scale_fill_hypso_c(
    name = TeX(r"(Number of month)"),
    palette = "wiki-schwarzwald-cont", 
    na.value = "transparent")
```

# Ensemble result

```{r}
#---- Prediction ----
y_m <- c('2019', '8') 
load(file = paste0(my.path, "/Models_ens/Ens1_err/ENS1_pred2_2019_08.Rdata"))
amaz.2019.08.ras <- rast(paste0(burntArea.path, '/burntarea_working_', y_m[1], '_', y_m[2], '.tif'))
pred.ras <- amaz.2019.08.ras
values(pred.ras) <- -1
tmp <- as.data.table(pred.ras, cell=T, xy=T)

# Prepare the value of each cell
prob.auc.amaz <- merge(x = as.data.table(prob.ens)[, c('cell', 'p1.auc')], y = tmp[, 'cell'], by = "cell", all = TRUE)

values(pred.ras) <- prob.auc.amaz$p1.auc
# writeRaster(pred.ras, paste0(path0, "/Models_ens/Ens1_err/ENS1_pred_2019_08.tif"), overwrite=F)

# plot

levels(amaz.2019.08.ras) <- data.frame(id=c(-2, 0, 1), val=c('-2', '0', '1'))
my.colors <- c("mediumblue", "mediumseagreen", "firebrick")
amaz.2019.08.m.ras <- amaz.2019.08.ras %>% mask(mask = amaz.basin.shp)

#---- Zones ----

v <- ext(amaz.basin.shp)
x.min <- v$xmin[[1]] - 1; x.max <- v$xmax[[1]] + 1; 
y.min <- v$ymin[[1]] - 1; y.max <- v$ymax[[1]] + 1; 

XY <- list(
  c(x.min, -0.84e+06, y.min, y.max),                       # z1
  c(-0.84e+06, +0.55e+06, 3.75e+06, y.max),                # z2
  c(-0.84e+06, +0.55e+06, 2.6e+06, 3.75e+06),              # z3
  c(-0.84e+06, -0.05e+06, 2.22e+06, 2.6e+06),              # z4
  c(-0.84e+06, -0.51e+06, y.min, 2.22e+06),                # z5
  c(-0.51e+06, -0.05e+06, y.min, 2.22e+06),                # z6
  c(-0.05e+06, 0.6e+06, y.min, 2.6e+06),                   # z7
  c(0.6e+06, 0.95e+06, y.min, 2.6e+06),                    # z8
  c(+0.55e+06, x.max, 3.24e+06, y.max),                    # z9
  c(+0.55e+06, x.max, 2.86e+06, 3.24e+06),                 # z10
  c(+0.55e+06, 0.95e+06, x.max, y.min, 2.6e+06, 2.86e+06)  # z11
)

nz <- length(XY)
zl <- list()
for (i in 1:(nz-1)){
  zl[[i]] <- rbind(
    c(XY[[i]][1], XY[[i]][3]), 
    c(XY[[i]][2], XY[[i]][3]), 
    c(XY[[i]][2], XY[[i]][4]), 
    c(XY[[i]][1], XY[[i]][4])
  ) %>% 
    cbind(object=i, part=1, ., hole=0) %>% 
    as.data.table()
}

i <- nz
zl[[i]] <- rbind(
  c(XY[[i]][2], XY[[i]][4]), 
  c(XY[[i]][3], XY[[i]][4]), 
  c(XY[[i]][3], XY[[i]][6]), 
  c(XY[[i]][1], XY[[i]][6]), 
  c(XY[[i]][1], XY[[i]][5]), 
  c(XY[[i]][2], XY[[i]][5])
) %>% 
  cbind(object=i, part=1, ., hole=0) %>% 
  as.data.table()

z <- rbindlist(zl, use.names=T) %>% as.matrix()
colnames(z)[3:4] <- c('x', 'y')

p <- vect(z, "polygons")
crs(p) <- crs(amaz.basin.shp)

#---- Plot ----
my.labels <-  c( 'z1',      'z2',      'z3',     'z4',      'z5',      'z6',     'z7',     'z8',     'z9',    'z10',    'z11')
x.coor <- c(-1.50e+06, -0.50e+06, -0.50e+06, -0.5e+06, -0.70e+06, -0.20e+06, 0.30e+06,  0.8e+06, 1.30e+06, 1.60e+06, 1.35e+06)
y.coor <- c( 3.05e+06,  4.05e+06,  3.05e+06,  2.4e+06,  1.85e+06,  1.85e+06, 1.90e+06, 2.00e+06, 3.75e+06, 3.05e+06, 2.50e+06)

p.zones <- myPlot(Q1, "Maximum value for each cell", theme=1) +
  geom_spatvector(data = p, fill = NA, color = "black") +
  coord_sf(datum = pull_crs(Q1)) + 
  scale_fill_manual(name = "Burnt Area", values = alpha(my.colors, 0.75), na.translate=FALSE) +
  annotate("text", label = my.labels, x = x.coor, y = y.coor, size = 6, colour = "black")
p.zones

#---- Plot ----
myPlot(
  pred.ras
  # , title = "Pattern for cells that may correspond to 'water' or \n'no fire' regions"
  # , paste0("Total number of cells = ", 
  #          formatC(length(cells(rand.p2.all.rast)), format="f", big.mark=",", digits=0) )
  # , theme = 1
) +
  geom_spatvector(data = p, fill = NA, color = "black") +
  scale_fill_whitebox_c(
    # name = TeX(r"(\overset{Evapotranspiration}{$\textit{(kg/m^2s)}$})"),
    name = TeX(r"(Probability)"),
    palette = "bl_yl_rd", 
    na.value = "transparent")

 myPlot(amaz.2019.08.m.ras, "Burnt Area in August 2019") +
  scale_fill_manual(name = "Burnt Area", values = my.colors, na.translate=FALSE)
```

```{r}
#---- Prepare data ----
# Burnt Area of the month
y_m <- c('2019', '8') 
amaz.2019.08.ras <- rast(paste0(burntArea.path, '/burntarea_working_', y_m[1], '_', y_m[2], '.tif'))
levels(amaz.2019.08.ras) <- data.frame(id=c(-2, 0, 1), val=c('-2', '0', '1'))
amaz.2019.08.m.ras <- amaz.2019.08.ras %>% mask(mask = amaz.basin.shp)
my.colors <- c("mediumblue", "mediumseagreen", "firebrick")
# The probability of the prediction
load(file = paste0(my.path, "/Models_ens/Ens1_err/ENS1_pred2_2019_08.Rdata"))
pred.ras <- amaz.2019.08.ras
values(pred.ras) <- -1
tmp <- as.data.table(pred.ras, cell=T, xy=T)
prob.auc.amaz <- merge(x = as.data.table(prob.ens)[, c('cell', 'p1.auc')], y = tmp[, 'cell'], by = "cell", all = TRUE)
values(pred.ras) <- prob.auc.amaz$p1.auc

# Zones SpatVector
zones.shp <- st_read("Amazon_new_data/1. Burnt Area/Zones_shp/Zones.shp")

#---- Plot ----
my.labels <-  c( 'z1',      'z2',      'z3',     'z4',      'z5',      'z6',     'z7',     'z8',     'z9',    'z10',    'z11')
x.coor <- c(-1.50e+06, -0.50e+06, -0.50e+06, -0.5e+06, -0.70e+06, -0.20e+06, 0.30e+06,  0.8e+06, 1.30e+06, 1.60e+06, 1.35e+06)
y.coor <- c( 3.05e+06,  4.05e+06,  3.05e+06,  2.4e+06,  1.85e+06,  1.85e+06, 1.90e+06, 2.00e+06, 3.75e+06, 3.05e+06, 2.50e+06)

myPlot(pred.ras, "Prediction of Burnt Area in August 2019", theme=1) +
  geom_spatvector(data = zones.shp, fill = NA, color = "gray50", linetype="dashed") + # solid, dashed, longdash, dotted, twodash, dotdash, blank
  coord_sf(datum = pull_crs(pred.ras)) + 
  scale_fill_hypso_c(
    # name = TeX(r"(\overset{Wind Speed }{$\textit{(m / s)}$})"),
    name = TeX(r"(Probability)"),
    palette = "gmt_globe_bathy", 
    trans = "pseudo_log",
    direction = -1,
    na.value = "transparent") +
  annotate("text", label = my.labels, x = x.coor, y = y.coor, size = 6, colour = "black")
```

```{r}
myPlot(
  pred.ras
  , title = "Prediction of Burnt Area in August 2019"
  # , paste0("Total number of cells = ", 
  #          formatC(length(cells(rand.p2.all.rast)), format="f", big.mark=",", digits=0) )
  # , theme = 1
) +
  scale_fill_hypso_c(
    # name = TeX(r"(\overset{Soil Moisture }{$\textit{(mm)}$})"),
    name = TeX(r"(Probability)"),
    palette = "wiki-schwarzwald-cont", 
    na.value = "transparent")
```

# Animation

```{r}
nbr_colors <- colorRampPalette(brewer.pal(9, "YlOrBr"))(1000)
# nbr_colors <- colorRamps::green2red(1000)
library(animation)
library(terra)
ras <- wind.rast[[1:20]]
animation::saveGIF(terra::animate(ras, col=nbr_colors),interval = 0.2, ani.width=800)



saveGIF({
  brownian.motion(pch = 21, cex = 5, col = "red", bg = "yellow")
}, movie.name = "brownian_motion.gif", interval = 0.1, nmax = 30, 
  ani.width = 600)
```

## gifski

### save rast as png

```{r}
predct.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Models_ens/Ens2_new/Predictions"
# import datra
predct.rast <- 
  list.files(paste0(predct.path, "/ras"), full.names=TRUE, pattern = ".tif$") %>% 
  rast()
# Rename layers
# precipitation.rast <- renameLayers(precipitation.rast, 'precipitation_working_', '')
# precipitation.rast <- precipitation.rast[[ordered.names]]
# Create `rts` object
# precipitation.rts <- rts(precipitation.rast, seq.dates)
# plot 


plot.name <- paste0(predct.path, "/png/Prediction_", names(ras), ".png")
png(filename = plot.name)


ras <- predct.rast[["2003_01"]]

plot(myPlot(ras, title = "Probability of fire", v_unit=c(3, 40, 3, 0)) +
  scale_fill_whitebox_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{Probability}$)"),
    palette = "muted", 
    breaks = c(0,0.2,0.4,0.6,0.8, 1),
    na.value = "transparent"))

plot(myPlot(ras, title = "Probability of fire", v_unit=c(3, 40, 3, 0)) +
  scale_fill_hypso_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{Probability}$)"),
    palette = "meyers_hypso", 
    na.value = "transparent"))

plot(myPlot(ras, title = "Probability of fire", v_unit=c(3, 40, 3, 0)) +
  scale_fill_hypso_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{Probability}$)"),
    palette = "arctic_hypso", 
    na.value = "transparent"))

plot(myPlot(ras, title = "Probability of fire", v_unit=c(3, 40, 3, 0)) +
  scale_fill_hypso_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{Probability}$)"),
    palette = "nordisk-familjebok_hypso", 
    na.value = "transparent"))

plot(myPlot(ras, title = "Probability of fire", v_unit=c(3, 40, 3, 0)) +
  scale_fill_whitebox_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{Probability}$)"),
    direction = -1,
    palette = "pi_y_g", 
    na.value = "transparent"))

plot(myPlot(ras, title = "Probability of fire", v_unit=c(3, 40, 3, 0)) +
  scale_fill_whitebox_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{Probability}$)"),
    # direction = -1,
    palette = "soft", 
    na.value = "transparent"))


dev.off()
```

### Real data

```{r}
 title=NULL
sub_title=NULL
theme=2
na_value=NULL
max_cell=1e6
x_angle=0
b_size=16
v_unit=c(2, 2, 2, 2) # c(t, r, b, l)

seq.m <- names(predct.rast)
rast <- burntArea.rast[[seq.m[2]]] %>% mask(amaz.basin.shp)
rast[rast < 0] <- NA
b_size <- 16
ggplot() +
  theme_bw(base_size=b_size) +
  geom_spatraster_contour(
    data = rast,
    aes(color = "red"),
    binwidth = 1,
  ) +
    geom_spatvector(data = amaz.basin.shp$geometry, fill = NA, color = "gray40") +
    scale_x_continuous(labels = function(x) format(x, scientific = T)) +
    scale_y_continuous(labels = function(x) format(x, scientific = T)) +
    ggtitle(label=title, subtitle=sub_title) +
    coord_sf(datum = pull_crs(rast)) +
    theme(
      axis.text.x = element_text(angle = x_angle)
      , plot.margin = unit(v_unit, "pt")
    ) 
```

```{r}
predct.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Models_ens/Ens2_new/Predictions"
# import datra
predct.rast <- 
  list.files(paste0(predct.path, "/ras"), full.names=TRUE, pattern = ".tif$") %>% 
  rast()
#
max_prb <- minmax(predct.rast) %>% max() %>% round(., digits = 1)
n.month <- "2007_01"
p.lst <- list()
for (n.month in seq.m){
  ba.rast <- burntArea.rast[[n.month]] %>% mask(amaz.basin.shp); ba.rast[ba.rast < 0] <- NA
  p.lst[[n.month]] <- myPlot(predct.rast[[n.month]], title = "Probability of fire in Amazon", sub_title = paste0("Month ", n.month), v_unit=c(3, 40, 3, 0)) +
    scale_fill_hypso_c(
      # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
      name = TeX(r"($\textit{Prob}$)"),
      palette = "arctic_hypso", 
      limits = c(0, max_prb),
      na.value = "transparent") +
    geom_spatraster_contour(
      data = ba.rast,
      aes(color = "red"),
      binwidth = 1,
      linewidth = 0.3
    ) +
    coord_sf(datum = pull_crs(predct.rast[[n.month]]))
}
p.lst
```

### Plot

```{r}
predct.path <- "/Users/ABIDM/Documents/Project/Code_official_data/Models_ens/Ens2_new/Predictions"
# import data
predct.rast <- 
  list.files(paste0(predct.path, "/ras"), full.names=TRUE, pattern = ".tif$") %>% 
  rast()
#
max_prb <- minmax(predct.rast) %>% max() %>% round(., digits = 1)
n.month <- "2010_01"
# Plot
title = "Fire in Amazon"
sub_title = paste0("Month ", n.month)
na_value = NULL
max_cell = 1e6
x_angle = 0
b_size = 16
v_unit = c(2, 2, 2, 2) # c(t, r, b, l)
rast <- predct.rast[[n.month]]
ba.rast <- burntArea.rast[[n.month]] % >% mask(amaz.basin.shp); ba.rast[ba.rast < 0] <- NA
ggplot() +
  geom_spatraster(data = rast, maxcell = max_cell) +
  geom_spatvector(data = amaz.basin.shp$geometry, fill = NA, color = "gray40") +
  geom_spatraster_contour(data = ba.rast, aes(col = "1"), binwidth = 1, linewidth = 0.3) +
  labs(col = TeX(r"($\textit{\overset{Fire}{events}})")) +
  scale_x_continuous(labels = function(x) format(x, scientific = T)) +
  scale_y_continuous(labels = function(x) format(x, scientific = T)) +
  ggtitle(label=title, subtitle=sub_title) +
  theme_bw(base_size=b_size) +
  theme(
    axis.text.x = element_text(angle = x_angle)
    , plot.margin = unit(v_unit, "pt")
  ) +
  scale_fill_hypso_c(
    # name = TeX(r"(\overset{Land Surface }{Temperature $\textit{(K)}$})"),
    name = TeX(r"($\textit{Probility}$)"),
    palette = "arctic_hypso", 
    limits = c(0, max_prb),
    na.value = "transparent") +
  coord_sf(datum = pull_crs(predct.rast[[n.month]]))
```

```{r}
# Create gif from temporary pngs
allfiles <- file.path(paste0(predct.path, "/png"), paste0("Prediction_",all_layers, ".png"))
gifski::gifski(allfiles,
  loop = TRUE, delay = 1,
  gif_file = "Amazon.gif",
  width = 1400, height = 865
)
```


```{r}
data("hypsometric_tints_db")
hypsometric_tints_db
```


```{r}
allfiles <- file.path(tempdir(), paste0(all_layers, ".png"))
gifski::gifski(allfiles,
  loop = TRUE, delay = 1 / 6,
  gif_file = "winter_2021.gif",
  width = 1600, height = 1200
)
```


# Variable Importance

## Initialize and Connect to H2O 

```{r}
library(h2o)
h2o.init(ip = 'localhost', port = 50001, nthreads = 6)
h2o.no_progress()
```

## Import models

```{r}
Models <- c('GLM', 'GBM', 'XGB', 'RF')
Metrics <- c('MSE', 'RMSE', 'LogLoss', 'Mean Per-Class Error', 'AUC', 'AUCPR', 'Gini', 'R^2')
nbr.metr <- length(Metrics)
nbr.m <- length(Models)
nbr.z <- 11
zone.lst <- c(1:nbr.z)
models.lst <- list()
for (zone in zone.lst){ # zone <- 1
  for (m in Models){ # m <- 'GBM'
    model.az <- h2o.upload_model(paste0(path0, "/Models/", m, "/", m, "_AZ", zone, "_model")) # 0.855 s
    models.lst <- append(models.lst, model.az)
  }
}

```





```{r}
h2o.removeAll()
h2o:::.h2o.garbageCollect()
h2o:::.h2o.garbageCollect()
h2o:::.h2o.garbageCollect()
h2o.ls()
# h2o.shutdown(prompt = FALSE)
```


















